### 2022-08-09 main/chapter_2ljets.tex
resolve todos for: asymptotic formulae ref, pdg rounding, claimed contributions, forward refs

% upper limits
% definitely want this after pictures
\FloatBarrier
\begin{table}[tp]
\centering
% custom separator for aligned +-
% https://stackoverflow.com/a/2132998
\begin{tabular*}{\textwidth}{lr@{$~\pm~$}lcccccc}
& \multicolumn{2}{c}{Fitted}
& Data
& $A\epsilon\sigma_{\mathrm{obs}}^{95}~\mathrm{fb}$
& $S_{\mathrm{obs}}^{95}$
& $S_{\mathrm{exp}}^{95}$
& $\clb$
& $p_{s\!=\!0}$
\\[1.5ex]
DR-OffShell & $22.1$ & $2.7$ & $21$ & $0.10$ & $14.3$ & $12.3^{+4.7}_{-3.1}$ & $0.68$ & $0.50$
\\[.5ex]
DR-Low & $22$ & $4$ & $18$ & $0.08$ & $10.8$ & $15.3^{+5.7}_{-4.0}$ & $0.09$ & $0.50$
\\[.5ex]
DR-Int & $35$ & $4$ & $38$ & $0.15$ & $20.9$ & $17.5^{+5.9}_{-3.9}$ & $0.73$ & $0.23$
\\[.5ex]
DR-High & $3.9$ & $0.5$ & $0$ & $0.02$ & $3.0$ & $5.6^{+2.2}_{-1.5}$ & $0$ & $0.50$
\\[.5ex]
DR-$\llbb$ & $0.51$ & $0.20$ & $0$ & $0.02$ & $3.0$ & $3.0^{+1.3}_{-0.0}$ & $0.19$ & $0.50$
\\[.5ex]
\end{tabular*}
\caption[
Upper limit results in $\twoljets$-electroweak discovery regions
]{%
Upper limit results in $\twoljets$-electroweak discovery
regions~\cite{atlas2022searches}.
The fitted yield is in the background-only model constrained by the data in each region.
Limits are intended to reflect constrains on additive signal contributions.
\\[0.5em]
Left to right:
the region name,
the post-fit background expectation,
the number of data observed,
the observed $95\%$ $\cls$ upper limit on the visible cross-section
$\langle\epsilon\sigma\rangle_\mathrm{obs}^{95}$,
its corresponding signal expectation $S_\mathrm{obs}^{95}$,
the expected $95\%$ upper limits on the signal expectation $S_\mathrm{exp}^{95}$
as would be obtained were the test statistic given by its central or
$\pm1$-sigma variations,
$\clb$ evaluated with the signal expectation set to the observed upper limit,
and the discovery $p$-value (capped at $0.5$) with its equivalent significance.
\\[0.5em]
Upper limits use the one-sided profile likelihood test statistic.
The discovery $p$-value uses a profile likelihood test statistic in a one-sided test.
All $p$-values are estimated by simulation of alternative data.
\\
\TODO{add labels and references to asymptotic formulae paper}
\\
\TODO{reference pdg rounding}
}
\label{tab:2ljets_discovery}
\end{table}

@@

\begin{itemize}
\item Design, implementation, and execution of the electroweak part of the analysis.
\item Service as `analysis contact' from January to October 2020.
\item Production of main data inputs for all three $\twoljets$ analyses
and the RJR $3\ell$ search~\cite{atlas_rjr_3l_SUSY_2019_09},
Production of all systematic variations for the background samples;
other group members assisted with the central background sample.
Production of all electroweak signal samples and their systematic variations.
\item \TODO{?Evaluation of Monte Carlo $Z$+jets uncertainties for RJR signal regions
which indicated they exceeded $100\%$}
\item \TODO{Various collaborations with group members on scripts to conduct
tasks which were common between the searches.}
\item Integration of the electroweak analysis with \atlas\ combinations and
pMSSM scan efforts, by performing their validation studies and serializing the
analysis results.
\item Preparation of electroweak results for publication in the
paper~\cite{atlas2022searches} and the public HEPData
record~\cite{maguire2017hepdata}.
\end{itemize}
Except where specified, I produced all figures presented in this thesis.
Many plotting scripts, however, are modified from versions inherited from
many \atlas\ members past and present, to whom I am grateful.

@@

In the context of $\twoljets$ selections, $\metsig$ suppresses $\ttbar$
contributions more than $\met$ alone, plausibly since $\ttbar$ events are
accompanied by more hadronic activity than other sources of $\ptmiss$.
What remains of the standard model in the high-$\metsig$ tail of $\twoljets$
selections is almost pure \diboson\ -- $ZZ$ with a hard $Z\rightarrow \nu\nu$,
and $W^\pm Z$ where the charged lepton in $W^\pm\rightarrow\ell^\pm\nu$
is not reconstructed.
\TODO{plot of \diboson\ decomposition here?}
These effects are demonstrated from simulations in
Figure~\ref{fig:2ljets_presel_met}.

@@

The $\twoljets$-electroweak search implements a total of $136$ modifiers in its
background-only model, and more for each signal model.
Of those I evaluated more manually, some of the more interesting or important
will be discussed in the remainder of this section ---
\TODO{update this}
Fake/non-prompt lepton backgrounds in Section~\ref{sec:2ljets_mm_fakes},
some alternative samples in Section~\ref{sec:2ljets_sample_comparisons},
various less impactful effects in
Section~\ref{sec:2ljets_other_background_variations},
and signal modelling uncertainties in
Section~\ref{sec:2ljets_signal_variations}.

>>>>

% upper limits
% definitely want this after pictures
\FloatBarrier
\begin{table}[tp]
\centering
% custom separator for aligned +-
% https://stackoverflow.com/a/2132998
\begin{tabular*}{\textwidth}{lr@{$~\pm~$}lcccccc}
& \multicolumn{2}{c}{Fitted}
& Data
& $A\epsilon\sigma_{\mathrm{obs}}^{95}~\mathrm{fb}$
& $S_{\mathrm{obs}}^{95}$
& $S_{\mathrm{exp}}^{95}$
& $\clb$
& $p_{s\!=\!0}$
\\[1.5ex]
DR-OffShell & $22.1$ & $2.7$ & $21$ & $0.10$ & $14.3$ & $12.3^{+4.7}_{-3.1}$ & $0.68$ & $0.50$
\\[.5ex]
DR-Low & $22$ & $4$ & $18$ & $0.08$ & $10.8$ & $15.3^{+5.7}_{-4.0}$ & $0.09$ & $0.50$
\\[.5ex]
DR-Int & $35$ & $4$ & $38$ & $0.15$ & $20.9$ & $17.5^{+5.9}_{-3.9}$ & $0.73$ & $0.23$
\\[.5ex]
DR-High & $3.9$ & $0.5$ & $0$ & $0.02$ & $3.0$ & $5.6^{+2.2}_{-1.5}$ & $0$ & $0.50$
\\[.5ex]
DR-$\llbb$ & $0.51$ & $0.20$ & $0$ & $0.02$ & $3.0$ & $3.0^{+1.3}_{-0.0}$ & $0.19$ & $0.50$
\\[.5ex]
\end{tabular*}
\caption[
Upper limit results in $\twoljets$-electroweak discovery regions
]{%
Upper limit results in $\twoljets$-electroweak discovery
regions~\cite{atlas2022searches}.
The fitted yield is in the background-only model constrained by the data in each region.
Limits are intended to reflect constrains on additive signal contributions.
\\[0.5em]
Left to right:
the region name,
the post-fit background expectation,
the number of data observed,
the observed $95\%$ $\cls$ upper limit on the visible cross-section
$\langle\epsilon\sigma\rangle_\mathrm{obs}^{95}$,
its corresponding signal expectation $S_\mathrm{obs}^{95}$,
the expected $95\%$ upper limits on the signal expectation $S_\mathrm{exp}^{95}$
as would be obtained were the test statistic given by its central or
$\pm1$-sigma variations,
$\clb$ evaluated with the signal expectation set to the observed upper limit,
and the discovery $p$-value (capped at $0.5$) with its equivalent significance.
\\[0.5em]
Upper limits use the one-sided profile likelihood test
statistic~\cite{Cowan:2010js}.
The discovery $p$-value uses a profile likelihood test statistic in a one-sided
test.
All $p$-values are estimated by simulation of alternative data.
Values with errors are rounded to one or two significant figures according
the the PDG rounding scheme~\cite{pdg2020review}.
}
\label{tab:2ljets_discovery}
\end{table}

@@

\begin{itemize}
\item Design, implementation, and execution of the electroweak part of the analysis.
\item Service as `analysis contact' from January to October 2020.
\item Production of main data inputs for all three $\twoljets$ analyses
and the RJR $3\ell$ search~\cite{atlas_rjr_3l_SUSY_2019_09},
Production of all systematic variations for the background samples;
other group members assisted with the central background sample.
Production of all electroweak signal samples and their systematic variations.
\item Evaluation of $Z$+jets uncertainties for RJR signal regions through
from simulations. These exceeded $100\%$ relative variation and supported the
use of data-driven methods.
\item Various collaborations with all other group members on common core tasks
that are shared between the searches of the $\twoljets$ project.
\item Integration of the electroweak analysis with \atlas\ combinations and
pMSSM scan efforts, by performing their validation studies and serializing the
analysis results.
\item Preparation of electroweak results for publication in the
paper~\cite{atlas2022searches} and the public HEPData
record~\cite{maguire2017hepdata}.
\end{itemize}
Except where specified, I produced all figures presented in this thesis.
Many plotting scripts, however, are modified from versions inherited from
many \atlas\ members past and present, to whom I am grateful.

@@

In the context of $\twoljets$ selections, $\metsig$ suppresses $\ttbar$
contributions more than $\met$ alone, plausibly since $\ttbar$ events are
accompanied by more hadronic activity than other sources of $\ptmiss$.
This effect in the $\metsig$ distribution is demonstrated in simulations in
Figure~\ref{fig:2ljets_presel_met}.
What remains of the standard model in the high-$\metsig$ tail of $\twoljets$
selections is almost pure \diboson\ -- $ZZ$ with a hard $Z\rightarrow \nu\nu$,
and $W^\pm Z$ where the charged lepton in $W^\pm\rightarrow\ell^\pm\nu$
is not reconstructed; these subcomponents of the \diboson\ process are discussed
later in Section~\ref{sec:2ljets_validation} and displayed in
Figure~\ref{fig:2ljets_splits_\diboson}.

@@

The $\twoljets$-electroweak search implements a total of $136$ modifiers in its
background-only model, and more for each signal model.
Of those not evaluated automatically, some of the more interesting or important
will be discussed in the remainder of this section.
Fake/non-prompt lepton backgrounds in Section~\ref{sec:2ljets_mm_fakes},
some alternative samples in Section~\ref{sec:2ljets_sample_comparisons},
various less impactful effects in
Section~\ref{sec:2ljets_other_background_variations},
and signal modelling uncertainties in
Section~\ref{sec:2ljets_signal_variations}.
